---
title: "{{ study_name }}"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<!-- README.md is generated from README.Rmd. Please edit that file -->

The data for the {{ study_name }} protocol can be found here: [insert path]. You can install the data package using devtools:

```{r install-example, eval=FALSE}
# if on github:
devtools::install_github(repo = "FredHutch/reponame.git")

# if on network drives:
devtools::install_git("path")
```

## SRAs and Study-Specific Training (SST)

| **SRA**| **STT Completed** | **Date**   |
|--------|-------------------|------------|
| Name   | No                | YYYY-MM-DD |


## Final Distributed Report Table

| **Assay Directory**| **Version** | **Date**   | **Unblinded** |
|--------------------|-------------|------------|---------------|
| [Assay](link)      | [0.0](link) | YYYY-MM-DD | Yes/No        |


## Analysis Notes

The protocol presentation can be found [here](link).

The lab study plan can be found [here](link).








