---
title: "Your Presentation Title"
subtitle: "A subtitle if you want one"
author: "Your Name"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  powerpoint_presentation:
    reference_doc: "`r system.file('templates', 'slides', 'Fred-Hutch-Gradient-Template.pptx', package = 'VISCtemplates')`"   # "Gradient" or "Light" or "Navy"
    slide_level: 2
---

<!--
NOTE ABOUT SLIDE BREAKS
   - A level-1 heading (#) starts a Section Header slide.
   - A level-2 heading (##) starts a new content slide (set via slide_level: 2 above).
   - A horizontal rule (---) also forces a new slide if you need one without a heading.
-->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(ggplot2)
```

# Section One

## Agenda

- Point one
- Point two
- Point three

## A Bulleted Content Slide

- First idea
  - Supporting detail
- Second idea
- Third idea with **bold** and *italics*

## A Plot Slide

```{r demo-plot, fig.width=7, fig.height=4}
ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point(size = 2) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "Weight (1000 lbs)", y = "Miles per gallon",
       title = "Fuel Efficiency vs. Weight") +
  theme_minimal(base_size = 14)
```

## A Table Slide

```{r demo-table}
knitr::kable(head(mtcars[, 1:5]), caption = "First rows of mtcars")
```

# Section Two

## Two-Content Layout

::: columns
::: {.column width="50%"}
**Left column**

- Item A
- Item B
:::
::: {.column width="50%"}
**Right column**

- Item C
- Item D
:::
:::

## Speaker Notes Example

This slide has notes attached below — they show up in Presenter View
in PowerPoint, and are stripped out of the beamer PDF.

::: notes
Remember to mention the Q3 numbers here and pause for questions.
:::

## Thank You

Questions?
