---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# VISCtemplates

The goal of VISCtemplates is to ...

## Installation

You can install the released version of VISCtemplates from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("VISCtemplates")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
## basic example code
```


## Spell Check

When the template is generated, it creates a list of custom words that are frequently ignored in VISC reports. This is saved in `inst/WORDLIST`. You can check spelling of files and update `WORDLIST` using the [`spelling` package](https://docs.ropensci.org/spelling/). The `spelling` package only spell checks text blocks, not code chunks.

```{r eval=FALSE}
install.packages("spelling")

library(spelling)

spell_check_files("BAMA/pt_report/BAMA_pt_report.Rmd")
```

