---
title: "Example VISC B-cell data analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Analysis of B-cell assays at VISC}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE}
# remotes::install_github("FredHutch/VISCfunctions", ref = "fas_ex_data")

library(VISCtemplates)
library(VISCfunctions)
library(tidyverse)
```

This vignette covers best practices at VISC for the analysis of B-cell data
from the flow cytometry and B-cell receptor (BCR) sequencing assays.

The example B-cell dataset we will use (which is made available through the
VISCfunctions package) consists of data from the IAVI G001 clinical trial.
The data has been re-processed from its original form to align with current
(as of May 2025) VISC standards for expected B-cell data schemas.

Here is a glimpse of the columns and example values in the dataset:

```{r load-example-data}
df_analysis <- G001_Bcell_flow_seq_PBMC # from VISCfunctions
glimpse(df_analysis)
```

And here is a summary of the unique endpoints in the dataset:

```{r}
df_analysis |>
  select(source_assay, endpoint, endpoint_value_type, bcell_population, percent_denominator) |>
  distinct() |>
  arrange(source_assay, endpoint) |>
  knitr::kable()
```

Note that this dataset includes both B-cell counts and perecentages.
Statistical analysis is typically done on the percentages, but the counts are
often reported as well - especially counts from the BCR sequencing assay.

# Response calls

## Response calls for BCF data

If you have a placebo group, can use this to estimate non-vaccine-induced variation over time:

```{r}

# use placebo participants to investigate non-vaccine-induced variation over time
df_placebo <- df_analysis |>
  filter(endpoint == "Percent of IgG+ B cells that are epitope-specific (KO-GT8++)",
         treatment == "DPBS sucrose") |>
  group_by(pubid) |>
  mutate(endpoint_value_baseline = endpoint_value[visit == -4]) |>
  ungroup() |>
  mutate(value_net = endpoint_value - endpoint_value_baseline) |>
  filter(visit != -4)

df_placebo |>
  select(pubid, visit, value_net) |>
  pivot_wider(names_from = visit, values_from = value_net)

# check distribution of the differences
df_placebo |>
  ggplot(aes(visitno, value_net, group = pubid)) +
  geom_line() +
  geom_point()

# seems reasonable to fit a normal distribution to these differences
mean_diff <- mean(df_placebo$value_net) # pretty close to zero
sd_diff <- sd(df_placebo$value_net)
cutoff <- qnorm(p = 0.99, mean = 0, sd = sd_diff)

df_response <- df_analysis |>
  filter(endpoint == "Percent of IgG+ B cells that are epitope-specific (KO-GT8++)") |>
  group_by(pubid) |>
  mutate(value_baseline = endpoint_value[visit == -4]) |>
  ungroup() |>
  mutate(value_net = endpoint_value - value_baseline) |>
  mutate(is_responder = value_net > cutoff)

df_response |>
  ggplot(aes(visitno, endpoint_value, shape = is_responder, group = pubid)) +
  geom_line(alpha = 0.5) +
  geom_point(alpha = 0.5) +
  scale_y_log10() +
  facet_wrap(~treatment)
```

If you don't have a placebo group, you can instead use variation across individuals
at baseline (rather than within-individual variation over time):

```{r}

df_baseline <- df_analysis |>
  filter(endpoint == "Percent of IgG+ B cells that are epitope-specific (KO-GT8++)",
         visit == -4)

df_baseline |>
  ggplot(aes(endpoint_value)) +
  geom_histogram(binwidth = 0.0005)

cutoff <- qnorm(p = 0.99,
                mean = mean(df_baseline$endpoint_value),
                sd = sd(df_baseline$endpoint_value))

df_response <- df_analysis |>
  filter(endpoint == "Percent of IgG+ B cells that are epitope-specific (KO-GT8++)") |>
  group_by(pubid) |>
  mutate(is_responder = endpoint_value > cutoff)

df_response |>
  ggplot(aes(visitno, endpoint_value, shape = is_responder, group = pubid)) +
  geom_line(alpha = 0.5) +
  geom_point(alpha = 0.5) +
  scale_y_log10() +
  facet_wrap(~treatment) +
  geom_hline(yintercept = cutoff, color = "red")
```

## Response calls for BCRseq data

For the VRC01-class endpoints, we do simple response calls, corresponding to
whether any VRC01-class sequences were observed after vaccination.

```{r}
df_responses_bcrseq <- df_analysis |>
  filter(bcell_population == "VRC01-class", endpoint_value_type == "percent") |>
  select(pubid, group, treatment, dose, visit, endpoint,
         bcell_population, percent_denominator, endpoint_value) |>
  distinct() |>
  mutate(response_call = endpoint_value > 0)

df_responses_bcrseq



df_analysis |>
  filter(endpoint_value_type == "count", source_assay == "sequencing", is.na(group) | visit == -4) |>
  select(pubid, visit, endpoint, endpoint_value) |>
  pivot_wider(names_from = "endpoint", values_from = "endpoint_value") |>
  arrange(visit) |>
  filter(!is.na(`Number of epitope-specific (KO-GT8++) IgG+ B cells that have BCR heavy and light chains sequenced`)) |>
  mutate(pct_vrc01_class_seq = 100 * `Number of epitope-specific (KO-GT8++) sequenced IgG BCRs that are VRC01-class` / `Number of epitope-specific (KO-GT8++) IgG+ B cells that have BCR heavy and light chains sequenced`) |>
  select(pct_vrc01_class_seq, `Number of epitope-specific (KO-GT8++) IgG+ B cells that have BCR heavy and light chains sequenced`)

chisq.test(matrix(c(1067986, 11694634, 285533, 3439404), 2, 2, byrow = TRUE))
```

# Graphical analysis

```{r}

library(patchwork)

plot_bcrseq_responses <- function(endpoint_name) {
  
  plt_df <- df_responses_bcrseq |>
    filter(endpoint == endpoint_name) |>
    mutate(dose = if_else(is.na(dose), 0, dose))
  
  plt_response_rate <- df_responses_bcrseq |>
    filter(!is.na(response_call)) |>
    group_by(treatment, dose, visit) |>
    summarize(response_rate = sum(response_call) / n()) |>
    ggplot(aes(as.factor(dose), response_rate)) +
    facet_wrap(~visit, nrow = 1, labeller = label_both) +
    geom_point(aes(color = treatment)) +
    theme_bw() +
    theme(legend.position = "none") +
    scale_y_continuous(labels = scales::percent, limits = c(0, 1))
  
  plt_magnitude <- plt_df |> 
    ggplot(aes(as.factor(dose), endpoint_value)) +
    facet_wrap(~visit, nrow = 1, labeller = label_both) +
    geom_boxplot(fill = NA) +
    geom_point(aes(color = treatment)) +
    theme_bw() +
    theme(legend.position = "none") +
    scale_y_sqrt(endpoint_name)
  
  plt_kinetics <- plt_df |> 
    ggplot(aes(as.factor(visit), endpoint_value, group = pubid)) +
    facet_wrap(~treatment, nrow = 1) +
    geom_line(alpha = 0.2) +
    geom_point(aes(color = treatment), alpha = 0.5) +
    scale_y_sqrt(endpoint_name) +
    theme_bw() +
    theme(legend.position = "bottom")
  
  plt_response_rate / plt_magnitude / plt_kinetics
  
}

plot_bcrseq_responses("Percent of IgG+ B cells detected as VRC01-class")
```



# Statistical comparisons

## Comparisons between groups (for each study visit)




## Comparisons between study visits (for each group)




