---
title: "{{ study_name}} Biological Endpoints"
output: github_document
---

## Terminology, abbreviations, and definitions of bnAb-class B cells

The following shorthand is used throughout this report:

- "IgG B cells" to mean IgG+ IgD- B cells as identified by flow cytometry,
- "IgD B cells" to mean IgG- IgD+ B cells as identified by flow cytometry,
- "Antigen-specific B cells" to mean ... B cells as identified
by flow cytometry,
- "Epitope-specific B cells" to mean ... B cells as identified by flow cytometry.

```{r bnab-class-definitions, tac.cap='bnAb class definitions used in this report.', results='asis'}

bnab_class_def_df <- data.frame(
  Epitope = c('CD4bs', 'CD4bs', 'CD4bs'),
  `bnab Class` = c('VRC01-class', 'IOMA-class', '8ANC131-class'),
  `Heavy Chain Features` = c('VH1-2', 'VH1-2', 'VH1-46'),
  `Light Chain Features` = c('5 AA CDR3', '7-9 AA CDR3', NA)) %>%
  setNames(c('Epitope', 'bnAb Class', 'Heavy Chain Features', 'Light Chain Features'))

knitr::kable(bnab_class_def_df)
```

## B-cell frequencies based on BCF data

The B-cell flow cytometry (BCF) endpoints covered by this report are detailed in
Table `r insert_ref("tab:flow-endpoints")`.
Each endpoint is a B-cell frequency that is calculated on a per-sample basis,
i.e., for each participant and study visit with data available.
Endpoint magnitudes are analyzed for each B-cell frequency, and positive
response calls and rates are computed and analyzed for B-cell frequencies out
of total IgG B
cells.^[Total IgG B cells were expected to be relatively stable over time,
unlike antigen-specific and/or epitope-specific B cells, which were
expected to be increase in frequency post-vaccination.]

```{r flow-endpoints}

df_example_flow_endpoints <- data.frame(
  endpoint = c("Percent of IgG B cells that are antigen-specific",
               "Percent of IgG B cells that are epitope-specific",
               "Percent of antigen-specific IgG B cells that are epitope-specific"),
  igx_type = rep("IgG", 3),
  antigen_specificity = rep("ANTIGEN", 3),
  epitope_specificity = c(NA, rep("EPITOPE", 2)),
  bnab_class = rep(NA, 3),
  percent_denominator = rep("IgG B cells", 3)
)

knitr::kable(df_example_flow_endpoints)
```

## B-cell frequencies based on BCRseq data

The primary B-cell receptor sequencing (BCRseq) endpoints covered by this report
are listed in Table `r insert_ref("tab:seq-endpoints")`.
Each of these is a B-cell frequency that is calculated on a per-sample basis,
i.e., for each participant and study visit with data available.

```{r seq-endpoints}

df_example_seq_endpoints <- data.frame(
  endpoint = c("Percent of epitope-specific IgG B cells that are VRC01-class",
               "Percent of epitope-specific IgG B cells that are IOMA-class"),
  igx_type = rep("IgG", 2),
  antigen_specificity = rep("ANTIGEN", 2),
  epitope_specificity = rep("CD4bs", 2),
  bnab_class = c("VRC01-class", "IOMA-class"),
  percent_denominator = rep("Epitope-specific IgG B cells", 2)
)

knitr::kable(df_example_seq_endpoints)
```

## Additional endpoints based on BCRseq data

This report also covers the following BCRseq endpoints, calculated for each
sequenced B cell and/or aggregated at the sample (participant and visit) level:

- Percent nucleotide-level somatic hypermutation in the V segment,
- Number of amino acid insertions and deletions,
- Presence of key VRC01-class residues,
- Distribution of germline V genes,
- CDR3 sequence length (number of amino acids),
- Clonotype (i.e., B-cell cluster) sizes and key characteristics of top clonotypes.

## B-cell frequencies based on combined BCF and BCRseq data

This report also covers certain endpoints based on combined BCF and BCRseq data,
which are detailed in Table `r insert_ref("tab:flow-seq-endpoints")`.
Each of these is a B-cell frequency that is calculated on a per-sample basis,
i.e., for each participant and study visit with data available.
The endpoint magnitudes are analyzed for each B-cell frequency, and positive response calls and rates are computed and analyzed for B-cell frequencies out of total IgG B 
cells.^[Total IgG B cells were expected to be relatively stable over time,
unlike antigen-specific and/or epitope-specific B cells, which were
expected to be increase in frequency post-vaccination.]

```{r flow-seq-endpoints}

df_example_flow_seq_endpoints <- data.frame(
  endpoint = c("Percent of IgG B cells that are VRC01-class",
               "Percent of antigen-specific IgG B cells that are VRC01-class"),
  igx_type = rep("IgG", 2),
  antigen_specificity = rep("ANTIGEN", 2),
  epitope_specificity = rep("CD4bs", 2),
  bnab_class = rep("VRC01-class", 2),
  percent_denominator = c("IgG B cells", "Antigen-specific IgG B cells")
)

knitr::kable(df_example_flow_seq_endpoints)
```
