---
title: "Your Presentation Title"
subtitle: "A subtitle if you want one"
author: "Your Name"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  powerpoint_presentation:
    reference_doc: template.pptx   # <- path to your custom PPT template
    slide_level: 2
  beamer_presentation:              # PDF slides (LaTeX-based); remove if not needed
    theme: "Madrid"
    slide_level: 2
---

<!--
HOW THIS FILE WORKS
====================
1. PPTX output:
   - `reference_doc` points to an existing .pptx file that defines your
     look and feel (fonts, colors, logo placeholder, slide master/layouts).
   - The easiest way to build that template: render this Rmd once WITHOUT
     reference_doc to get a default deck, open it in PowerPoint, edit the
     slide master / layouts to match your branding, save it as template.pptx,
     then point reference_doc at it. rmarkdown will map heading levels and
     content onto the layouts you defined (Title Slide, Title and Content,
     Two Content, etc.) based on their names in the template.
   - Put template.pptx in the same folder as this .Rmd (or use a full path).

2. PDF output:
   - `beamer_presentation` needs a LaTeX installation (e.g. TinyTeX:
     `tinytex::install_tinytex()`). It does NOT use the PowerPoint template;
     it has its own themes (Madrid, Berlin, Warsaw, etc.) and color themes
     (`colortheme: "dolphin"`).
   - If you specifically need a PDF that matches your PPTX template's look,
     the practical route is: render the PPTX above, then convert it to PDF
     with LibreOffice from the command line:
       soffice --headless --convert-to pdf your_slides.pptx
     LaTeX beamer and PowerPoint styling are different engines and can't
     share a template directly.

3. RENDERING
   - Render just the PPTX:
       rmarkdown::render("slides_template.Rmd", output_format = "powerpoint_presentation")
   - Render just the PDF:
       rmarkdown::render("slides_template.Rmd", output_format = "beamer_presentation")
   - Render everything listed in the YAML:
       rmarkdown::render("slides_template.Rmd", output_format = "all")

4. SLIDE BREAKS
   - A level-1 heading (#) starts a Section Header slide.
   - A level-2 heading (##) starts a new content slide (set via slide_level: 2 above).
   - A horizontal rule (---) also forces a new slide if you need one without a heading.
-->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(ggplot2)
```

# Section One

## Agenda

- Point one
- Point two
- Point three

## A Bulleted Content Slide

- First idea
  - Supporting detail
- Second idea
- Third idea with **bold** and *italics*

## A Plot Slide

```{r demo-plot, fig.width=7, fig.height=4}
ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point(size = 2) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "Weight (1000 lbs)", y = "Miles per gallon",
       title = "Fuel Efficiency vs. Weight") +
  theme_minimal(base_size = 14)
```

## A Table Slide

```{r demo-table}
knitr::kable(head(mtcars[, 1:5]), caption = "First rows of mtcars")
```

# Section Two

## Two-Content Layout

::: columns
::: {.column width="50%"}
**Left column**

- Item A
- Item B
:::
::: {.column width="50%"}
**Right column**

- Item C
- Item D
:::
:::

## Speaker Notes Example

This slide has notes attached below — they show up in Presenter View
in PowerPoint, and are stripped out of the beamer PDF.

::: notes
Remember to mention the Q3 numbers here and pause for questions.
:::

## Thank You

Questions?
