---
title: "{{ study_name }}"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<!-- README.md is generated from README.Rmd. Please edit that file -->

The data for the {{ study_name }} protocol can be found here: [insert path]. You can install the data package using the `remotes::install_github` or `VISCtemplates::install_git_workaround`:

```{r install-example, eval=FALSE}
# if on github:
remotes::install_github(repo = "FredHutch/reponame.git")

# if on network drives:
VISCtemplates::install_git_workaround("path/reponame.git")
```

## SRAs and Study-Specific Training (SST)

| **SRA**| **STT Completed** | **Date**   |
|--------|-------------------|------------|
| Name   | No                | YYYY-MM-DD |


## Final Distributed Report Table

| **Assay Directory**| **Version** | **Date**   | **Unblinded** |
|--------------------|-------------|------------|---------------|
| [Assay](link)      | [0.0](link) | YYYY-MM-DD | Yes/No        |

## Study Start Up Checklist

The SRA or designee is responsible for the following:

- [ ] Create TMF folder on network drive (SOP-0087 SCHARP Trial Master File)
- [ ] Create subfolders based on `J:\QUALITY\Quality - all access\15. eTMF Information\[Study ID - folder structure to copy]\`
  - [ ] Create `TMF/01_Trial_Management`
  - [ ] Create `TMF/11_Statistics`
- [ ] Download and fill out FRM-0086 Study Staff Roster
  - [ ] Change Department to DA. Start date is study assignment. Start all staff as blinded until unblinding information is received. 
  - [ ] Include: lead statistician, lead SRA, lead STP, PM, plus VISC leadership (VISC associate director, VISC program manager, statistical manager, VISC PI, and VISC CAVD co-PI).
  - [ ] Save the study roster as a Word doc (no date in name) AND as a certified PDF with the date appended to the file name in: `TMF/01.02.01 Trial Team Details/Study Team Staff Roster`
  - [ ] If blinding status changes or people join/leave the team, update the Word doc and save the new certified PDF with the new date appended in: `TMF/01.02.01 Trial Team Details/Study Team Staff Roster`
- [ ] Save a certified PDF of study assignment documentation in `TMF/01.02.01 Trial Team Details/Trial Team Documentation`
- [ ] Fill out FRM-0094 Statistical Team Roster and Study Blinding Access Control Form (SOP-0075 Study Blinding and Unblinding). It should match FRM-0086. Save a certified PDF copy to `11 Statistics\11.02 Randomization\11.02.06 End of Trial or Interim Unblinding - restricted\Statistical Team Roster and Study Blinding Access Control Doc`
- [ ] Create the folder `cavdNNN_closed` and add permissions based on FRM-0094. This is the LDAP group (template in SOP-0075 Appendix A)
  - [ ] On FRM-0094 specify creation of LDAP group `cavdNNN_closed` or `cavd_idNNN_closed` (all study team members, name depends on CAVD vs GHDC) and optionally `cavdNNN_rand` or `cavd_idNNN_rand` (only study team members who need access to unblinded treatment assignments) and list members of both.
  - [ ] Can create LDAP group by changing permission to the folders in command line (see VISC-Documentation for instructions) or
  - [ ] Can create LDAP group by emailing support@scharp.org
  - [ ] Ensure all unblinding data is only accessible to those in the LDAP
- [ ] Create analysis repository in github using VISCtemplates
  - [ ] Add staff and corresponding permissions to repo according to FRM-0086
- [ ] Sign, certify and save study specific individual (FRM-0116) or group (FRM-0117) training document to `01 Trial Management\01.04 Meetings\01.04.04 Trial Team Evidence of Training\Study Specific Training Affidavit\` (SOP-0002)
  - [ ] Using naming convention "DDMMMYYYY_NETWORK####_FileName_v1.0.pdf" (SOP-0007)

For details on documenting changes to the lead SRA, lead STP, or for more information see: https://github.com/FredHutch/VISC-Documentation/blob/main/Assay_Analyses/Cross-Assay-Instructions/Study_Startup/Study-Start-Up-SRA.md

## Analysis Notes

The protocol presentation can be found [here](link).

The lab study plan can be found [here](link).








