---
title: "{{ study_name }}"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The data for the {{ study_name }} protocol can be found here: [insert path]. You can install the data package using the `remotes::install_github` or `VISCtemplates::install_git_workaround`:

```{r install-example, eval=FALSE}
# if on github:
remotes::install_github(repo = "FredHutch/reponame.git")

# if on network drives:
VISCtemplates::install_git_workaround("path/reponame.git")
```

## Final Distributed Report Table

| **Assay Directory**| **Version** | **Date**   | **Unblinded** |
|--------------------|-------------|------------|---------------|
| [Assay](link)      | [0.0](link) | YYYY-MM-DD | Yes/No        |


## Analysis Notes

The protocol presentation can be found [here](link).

The lab study plan can be found [here](link).


## Study Documentation

### SRAs and Study-Specific Training (SST)

| **SRA**| **STT Completed** | **Date**   |
|--------|-------------------|------------|
| Name   | No                | YYYY-MM-DD |

### Study Start Up Documentation

All start up study documentation was completed by [SRA or lead analyst] on [YYYY-MM-DD]. 
[See checklist](Startup-Checklist.md).
